---
title: "Heteroscedasticity"
output: html_notebook
---

One of the assumptions for an ordinary least squares regression is that the errors
are constant across the domain of the model (variance does not vary with input).
This is known as homoscedasticity, and its counterpart is heteroscedasticity.

As homoscedasticity is an assumption of the formulation of OLS, if the model has
heteroscedasticity the model's hypothesis test are no longer valid, and the
estimators are not BLUE (Best Linear Unbiased Estimator).

## Bartlett Test

Used to test if variances across samples is equal.
Its null hypothesis is that the variances are equivalent.
$$ H_0: \sigma_1^2 = \sigma_2^2 = ... = \sigma_k^2 $$
Its alternative hypothesis is that at least one pair of variances are different.
$$ H_a: \sigma_i^2 \neq \sigma_j^2 \ \ \ \text{for at least one pair} (i,j) $$

```{r}
library(olsrr)
ols_test_bartlett(hsb, 'read', group_var = 'female')
```
## Breusch Pagan Test

Introduced by Trevor Breusch and Adrian Pagan in 1979. It tests whether the
variance of the errors from a regression is dependent on the values of the
independent variables by means of a $\chi^2$ test.

$$ H_0: \sigma_1^2 = \sigma_2^2 = ... = \sigma_k^2 $$
$$ H_a: \sigma_i^2 \neq \sigma_j^2 \ \ \ \text{for at least one pair} (i,j) $$
### Using fitted values
```{r}
model = lm(mpg ~ disp + hp + wt + drat, data = mtcars)
ols_test_breusch_pagan(model)
```

### Using independent variables
```{r}
model = lm(mpg ~ disp + hp + wt + drat, data = mtcars)
ols_test_breusch_pagan(model, rhs = TRUE)
```

### Using independent variables and performing multiple tests
```{r}
model = lm(mpg ~ disp + hp + wt + drat, data = mtcars)
ols_test_breusch_pagan(model, rhs = TRUE, multiple = TRUE)
```
The p-value can be adjusted with the `p.adj` parameter. Bonferroni, Sidak and
Holm's p-value adjustment are available.

## Score Test

Test for heteroscedasticity under the assumption that the errors are iid.

### Using fitted values
```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_score(model)
```

### Using independent variables
```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_score(model, rhs = TRUE)
```

### Specify variables
```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_score(model, vars = c('disp', 'hp'))
```

## F Test

F test for heteroscedasticity under the assumption that the errors are iid.

### Using fitted values
```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_f(model)
```

### Using independent variables
```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_f(model, rhs = TRUE)
```

### Specify variables

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_f(model, vars = c('disp', 'hp'))
```


## References
https://olsrr.rsquaredacademy.com/articles/heteroskedasticity.html