---
title: "Collinearity Diagnostics, Model Fit & Variable Contribution"
output: html_notebook
---

# Collinearity Diagnostics

Collinearity implies two variables are near perfect linear combinations of one
another. Multicollinearity involves more than two variables. In the presence of
multicollinearity, regression estimates are unstable and have high standard
errors.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_coll_diag(model)
```


## Variance Inflation Factors

They measure the inflation in the variances of the parameter estimates due to
collinearities that exist among the featuers. It is a measure of how much the
variance of the regression coefficient $\beta_k$ is inflated by the existence
of correlation among the features in the model.

A VIF of 1 means that there is no correlation. The general rule of thumb is that
VIFs exceeding 4 warrant further investigation. The steps to calculate VIF are:
*   Regress the $k^{th}$ feature on rest of the features in the model.
*   Compute $R_k^2$

$$ VIF = \frac{1}{1-R_k^2} = \frac{1}{Tolerance} $$
```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_vif_tol(model)
```

## Condition Index

Most multivariate statistical approaches involve decomposing a correlation
matrix into linear combinations of variables. The linear combinations are chosen
so that the first combination has the largest possible variance (subject to some
restrictions we won’t discuss), the second combination has the next largest
variance, subject to being uncorrelated with the first, the third has the
largest possible variance, subject to being uncorrelated with the first and
second, and so forth. The variance of each of these linear combinations is
called an eigenvalue. Collinearity is spotted by finding 2 or more variables
that have large proportions of variance (.50 or more) that correspond to large
condition indices. A rule of thumb is to label as large those condition indices
in the range of 30 or larger.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_eigen_cindex(model)
```

# Model Fit Assessment

## Residual Fit Spread Plot

Plot to detect non-linearity, influential observations and outliers. Consists of
side-by-side quantile plots of the centered fit and the residuals.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_fit_spread(model)
```

## Part & Partial Correlations

### Correlations
How much each feature uniquely contributes to $R^2$.

### Zero order
Pearson correlation coefficient between the dependent variable and the features.

### Part
Unique contribution of feature. How much the $R^2$ would decrease if the feature
was removed from the model.

### Partial
How much of the output variance is estimaited by the feature.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_correlations(model)
```

## Observed vs Predicted Plot

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_obs_fit(model)
```

## Lack of Fit F Test

The residual sum of squares can be decomposed into 2 components:
*   Due to lack of fit
*   Due to random variation

If most of the error is due to lack of fit and not just random error, the model
should be discarded and a new model must be built. The test assesses how much
of the error in the prediction is due to the lack of model fit.
```{r}
model = lm(mpg ~ disp, data = mtcars)
ols_pure_error_anova(model)
```

## Diagnostics Panel

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_diagnostics(model)

```

# Variable Contributions

## Residual vs Regressor Plots

Graph to determine whether we should add a new predictor to the model already
containing other predictors. The residuals from the model is regressed on the
new predictor and if the plot shows non random pattern, you should consider
adding the new predictor to the model.

```{r}
model = lm(mpg ~ disp + hp + wt, data = mtcars)
ols_plot_resid_regressor(model, 'drat')
```

## Added Variable Plot

Provides information about the marginal importance of a feature X, given the
other features already in the model.

Steps to construct an added variable plot:
*   Regress Y on all features other than X and store the residuals.
*   Regress X on all the other features in the model.
*   Construct a scatter plot with the residuals of the previous steps.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_added_variable(model)
```

## Residual Plus Component Plot

Steps to construct the plot:
*   Regress Y on all features including X.
*   Multiply $e$ with regression coefficient of X.
*   Construct plot of $eX$ and $X$.

The residual plus component plot indicates whether any non-linearity is present
in the relationship between Y and X and can suggest possible transformations for
linearizing the data.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_comp_plus_resid(model)
```



# References
https://olsrr.rsquaredacademy.com/articles/regression_diagnostics.html