---
title: "Measures of Influence"
output: html_notebook
---

It is possible for a single observation to have a great influence on the results
of a regression. It is, therefore, important to detect influential observations
to take them into consideration when interpreting results.

## Cook's D Bar Plot

Cook's distance was introduced in 1977 by American statician R Dennis Cook. It
depends on both the residual and leverage. The steps to compute it are:
*   delete observations one at a time
*   refit the regression model on the remaining observations
*   examine how much the fitted values changed after the observation was deleted
```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_cooksd_bar(model)
```

## Cook's D Chart
```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_cooksd_chart(model)
```
## DFBETAs Panel

(DF -> Degrees of Freedom)

DFBETA measures the difference in each parameter estimate with and without the
influential point. There is a DFBETA for each data point.

```{r}
model = lm(mpg ~ disp + hp + wt, data = mtcars)
ols_plot_dfbetas(model)
```

## DFFITS Plot

Proposed by Welsch and Kuh (1977). It is the difference between the $i^{th}$
fitted value obtained from the full data and the $i^{th}$ observation. Used to
identify influential data points.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_dffits(model)
```

## Studentized Residual Plot

Studentized deleted residuals is the deleted residual divided by its estimated
standard deviation. If an observation has a studentized deleted residual larger
than 3 (absolute value) we call it an outlier.
```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_stud(model)
```

## Standardized Residual Chart

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_stand(model)
```

## Studentized Residuals vs Leverage Plot


```{r}
model = lm(read ~ write + math + science, data = hsb)
ols_plot_resid_lev(model)
```

## Deleted Studentized Residual vs Fitted Values Plot

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_stud_fit(model)
```

## Hadi Plot

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_hadi(model)
```

## Potential Residual Plot

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_pot(model)
```


## Reference
https://olsrr.rsquaredacademy.com/articles/influence_measures.html



