---
title: "Residual Diagnostics"
output: html_notebook
---

## Residual QQ Plot

Used to violation of normality assumption.

```{r}
library(olsrr)
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_qq(model)
```

## Residual Normality Test

Tests for normality by means of multiple statistical tests.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_normality(model)
```

Correlation between observed residuals and expected residuals under normality.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_correlation(model)
```

## Residual vs Fitted Values Plot

A scatter plot of residuals vs fitted values to detect non-linearity, unequal
error variances and outliers.

```{r}
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_fit(model)
```

A good residual vs fitted plot has:
*   Residuals randomly spread around the 0 line.
*   Residuals forming a band around the 0 (like a pipe) indicating homogeneity
of error variance.
*   No residual visibly away from the random pattern.

## Residual Histogram

Used to detect violation of normality assumption.

```{r}
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_hist(model)
```


## References
https://olsrr.rsquaredacademy.com/articles/residual_diagnostics.html
