---
title: "Variable Selection Methods"
output: html_notebook
---

## All Possible Regression

For K potential features there are $2^k$ distinct possible regression models
to be tested.

```{r}
library(olsrr)
model = lm(mpg ~ disp + hp + wt + qsec, data =mtcars)
k = ols_step_all_possible(model)
k
```

`plot` can be used to show the panel of fit criteria for all possible regressions.
```{r}
plot(k)
```

## Best Subset Regression

Selects the best subset that perform the best at some objective criteria, such
as $R^2$, MS2, AIC, etc.

```{r}
model = lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
k = ols_step_best_subset(model)
k
plot(k)
```

## Stepwise Forward Regression

Buils a regression model from a set of features by entering them one by one
based on p values. Further details can be obtained by setting `details = true`.

```{r}
model = lm(y ~ ., data = surgical)
k = ols_step_forward_p(model)
k
plot(k)
```

## Stepwise Backward Regression

Builds a regression model starting with all features and removing them one
by one based on p values. Further details can be obtained by setting 
`details = true`.

```{r}
model = lm(y ~ ., data = surgical)
k = ols_step_backward_p(model)
k
plot(k)
```

## Stepwise regression

An iterative method in which features are entered and removed one by one and
decided upon by performing f or t tests to know if they were significant.

```{r}
model = lm(y ~ ., data = surgical)
k = ols_step_both_p(model)
k
plot(k)
```

## Stepwise AIC Forward Regression

An iterative method in which features are added one by one and decided upon by
checking the Akaike Information Criterion (less is better).

```{r}
model = lm(y ~ ., data =surgical)
k = ols_step_forward_aic(model)
k
plot(k)
```


## Stepwise AIC Backward Regression

An iterative method in which features are removed one by one and
decided upon by checking the Akaike Information Criterion (less is better).
```{r}
model = lm(y ~ ., data =surgical)
k = ols_step_backward_aic(model)
k
plot(k)
```

## Stepwise AIC Regression

An iterative method in which features are added and removed one by one, and
decided upon by checking the Akaike Information Criterion (less is better).

```{r}
model = lm(y ~ ., data = surgical)
k = ols_step_both_aic(model)
k
plot(k)
```

## Reference

https://olsrr.rsquaredacademy.com/articles/variable_selection.html


